---
output:
  pdf_document: default
  html_document: default
  word_document: default
---


```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, echo=FALSE, include=FALSE, results=FALSE}
library("readr") 
library("readxl") 
library("tidyverse")
library("dplyr")
library("dplyr")
library("ggplot2")
library("e1071")
library("tinytex")
```

#TASK 1 : Prescriptive Part

###Question 1 :
### I copied all my code from the last homework here
```{r}
tkurra_beans <- read_excel("Desktop/tkurra_beans.xlsx")
new_tkurrabeans <- tkurra_beans %>%
  select(-AspectRatio, -Area) %>%
  rename ("LengthMaj" = MajorAxisLength, "LengthMin" = MinorAxisLength) %>%
  filter(tkurra_beans$Class %in% c("SEKER", "DERMASON", "CALI", "BARBUNYA", "HOROZ")) %>%
  select(everything()) %>% 
  mutate(average_MinMax = rowMeans(select(., c("LengthMaj", "LengthMin")), na.rm = TRUE)) %>%
  mutate(Solidity_category = case_when(
    Solidity > 0.9889893 ~ "Largest category",
    Solidity > 0.9860581 & Solidity <= 0.9889893 ~ "Middle category",
    TRUE ~ "Lowest category")) %>%
  arrange(desc(Perimeter))

  head(new_tkurrabeans)
  view(new_tkurrabeans)

```


###Question 2 :
##I created a matrix by using the "cor" function to join all the variables the homework asks for including perimeter, eccentricity, convexarea, equivdiameter, and extent. Then I displayed the first couple of observations using the head function.
```{r}
beans_matrix <- cor(data.frame(new_tkurrabeans$Perimeter, new_tkurrabeans$Eccentricity, new_tkurrabeans$ConvexArea, new_tkurrabeans$EquivDiameter, new_tkurrabeans$Extent))

head(beans_matrix)
```

###Question 3 :
###I created a plot to compare the different classes and their solidity levels. I used the new_tkurrabeans dataset and made the x value the Classes variable and the y the Solidity. Then the fill I made the categories of the solidity variable.
```{r}
ggplot(new_tkurrabeans, aes(x = Class, y = Solidity, fill = Solidity_category)) +
  geom_bar(stat = "identity") +
  labs(title = "Comparison of Beans Solidity Distribution Across Classes",
       x = "Class",
       y = "Solidity",
       fill = "Solidity_category")

```

###Question 4 :
###I created a plot to make a histogram of the solidity variable in the newtkurra_beans dataset and added a density plot on top. I altered the alpha and colors to make it visually appealing and presentable as well. 
```{r}
ggplot(new_tkurrabeans, aes(x = Solidity)) +
  geom_histogram(aes(y = ..density..),      
                 bins = 10,               
                 fill = "skyblue",          
                 color = "black",        
                 alpha = 0.7) +            
  geom_density(color = "red", alpha = 0.7) + 
  labs(title = "Histogram with Bean's Density Overlay of Solidity") +  
  xlab("Solidity") +                      
  ylab("Density")                         
```

###Question 5 :
###I used what I did in the question before and used the facet_wrap function to create multiple plots of each class.
```{r}
ggplot(new_tkurrabeans, aes(x = Solidity)) +
  geom_histogram(aes(y = ..density..),  
                 bins = 10,            
                 fill = "skyblue",         
                 color = "black",        
                 alpha = 0.7) +          
  geom_density(color = "red", alpha = 0.7) + 
  labs(title = "Histogram with Bean's Density Overlay of Solidity in Different Classes") + 
  xlab("Solidity") +       
  ylab("Density") +                        
  facet_wrap(~ Class)  
```

#TASK 2 : Open-Ended Part

####<https://www.kaggle.com/datasets/jessemostipak/hotel-booking-demand> This dataset seemed intersting to me because I think it is interesting how during certain times of the year there are definelty more sales and reservations in hotel rooms. I wanted to use this dataset to be able to analyze how bookings may vary throughout the year. I also liked how it has many quantitative and qualitative variables to choose and analyze from.

```{r}
hotel_bookings <- read_csv("Desktop/hotel_bookings.csv")
View(hotel_bookings)
```



####Questions : 1. Does a certain hotel style have less cancellations than the other?
####2. Is the average stay in during weeknights greater in summer months than fall? 
```{r}
cancellation_summary <- hotel_bookings %>%
  group_by(hotel) %>%
  summarise(is_canceled = sum(is_canceled))
cancellation_summary
```

```{r}
summer_data <- subset(hotel_bookings$stays_in_week_nights, hotel_bookings$arrival_date_month %in% c("June", "July", "August"))
fall_data <- subset(hotel_bookings$stays_in_week_nights, hotel_bookings$arrival_date_month %in% c("September", "October", "November"))
avg_stay_summer <- mean(summer_data, na.rm = TRUE)
avg_stay_fall <- mean(fall_data, na.rm = TRUE)
avg_stay_fall
avg_stay_summer
```


```{r}



mean_data <- aggregate(is_canceled ~ hotel, data = hotel_bookings, FUN = mean)


ggplot(mean_data, aes(x = hotel, y = is_canceled)) +
  geom_bar(stat = "identity", fill = "beige", color = "tan", size = 2) +
  labs(title = "Average Cancelations for Hotels", x = "Hotels", y = "Average Caneleations")

```
```{r}
summer_data <- subset(hotel_bookings$stays_in_week_nights, hotel_bookings$arrival_date_month %in% c("June", "July", "August"))
fall_data <- subset(hotel_bookings$stays_in_week_nights, hotel_bookings$arrival_date_month %in% c("September", "October", "November"))

avg_stay_summer <- mean(summer_data, na.rm = TRUE)
avg_stay_fall <- mean(fall_data, na.rm = TRUE)

average_stay <- data.frame(
  Month = c("Summer", "Fall"),
  Average_Stay = c(avg_stay_summer, avg_stay_fall)
)

ggplot(average_stay, aes(x = Month, y = Average_Stay, fill = Month)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Stay During Weeknights",
       x = "Month",
       y = "Average Stay",
       fill = "Month")

```

####The summaries show that on average, there tends to be more cancelations in city hotels rather than resort hotels, which makes sense because resort hotels tend to be more expensive than city hotels. The other summary shows that the average stay in higher during weeknights is higher during the summer rather than fall. This also makes sense as that is when break happens and there is more free time compared to fall. More vacations tend to happen during the summer as well.